---
title: "EDA"
author: "Nisse Hermsen"
date: "`r Sys.Date()`"
output: html_document
---

# Bike sharing
url: https://archive.ics.uci.edu/dataset/275/bike+sharing+dataset

Import the required dependencies.
```{r import}
library(tidyverse)
library(fastDummies)
```
Proceed to read the data.
```{r read bike}
bike_df <- read.csv("./data/bike.csv", header = TRUE, sep = ",")

head(bike_df, 20)
```

# Preprocessing
TODO:
- Categorical data transformation
- Missing data
- Distributions

We first factorise the categorical data present within the data set. A method to do this for regression analysis is by
creating dummy variables (a.k.a. dummy coding). By removing the first dummy variable - leaving us with n-1 dummy
variables - we avoid multi-collinearity issues in  (https://cran.r-project.org/web/packages/fastDummies/fastDummies.pdf).\\
More info on categorical data in regression (https://stats.oarc.ucla.edu/spss/faq/coding-systems-for-categorical-variables-in-regression-analysis-2/#:~:text=Categorical%20variables%20require%20special%20attention,entered%20into%20the%20regression%20model.)

Further research: generalised linear model, using poisson distribution.
```{r factorise}
categorical_cols <- c('season', 'yr', 'mnth', 'weekday', 'weathersit')
new_df <- dummy_cols(bike_df, select_columns = categorical_cols, remove_first_dummy = TRUE)
head(new_df, 20)
```