---
title: "EDA"
author: "Aga, Karla, Nisse, Ole"
date: "`r Sys.Date()`"
output: html_document
---

# Bike sharing
## Load data
Data set is acquired from [machine-learning repo](https://archive.ics.uci.edu/dataset/275/bike+sharing+dataset).

Import the required dependencies.
```{r import, message = FALSE}
library(tidyverse)
library(fastDummies)
library(kableExtra)
```
Proceed to read the data.
```{r read bike}
source <- read.csv("../data/hour.csv", header = TRUE, sep = ",") %>%
    as_tibble()
```

## Preprocessing
Sub-select variables for linear model:
```{r select}
data <- select(source, dteday, hr, holiday, temp, atemp, hum, windspeed, cnt)
```

### Missing data
Since all of our current predictor variables are numeric (rather than text), we  
can first find missing values by checking for the standard `NA` notation. In this  
case, we do *not* find any `NA` values.
```{r missing standard}
anyNA(data)
```
That said, missing values can still be represented as numeric values (e.g.  
`0`, `999`, `inf`). Therefore we check for non-standard formatted missing values by  
looking at the distributions of the predictor values.  
```{r missing non-standard}
# hr
ggplot(data, aes(x = hr)) + 
    geom_histogram(binwidth = 1)

# temp
ggplot(data, aes(x = temp)) + 
    geom_histogram()

# atemp
ggplot(data, aes(x = atemp)) + 
    geom_histogram()

# windspeed
ggplot(data, aes(x = windspeed)) + 
    geom_histogram()

# hum
ggplot(data, aes(x = hum)) + 
    geom_histogram()

# cnt
ggplot(data, aes(x = cnt)) + 
    geom_histogram()
```

### Factorise hour data
Since hour is categorical, we need to factorise the vector data.
```{r segment hour}
# Increment the `hr` column by one for ease of indexing purposes
data$hr <- data$hr + 1

# Factorise the `hr` column using 4 categories / bins.
data <- data %>%
    mutate(hr_bin = cut(hr,
        breaks = c(0, 6, 12, 18, 24),
        labels = NULL
    ))

# Create a `mode` function for aggregation.
mode <- function(x) {
    which.max(table(x))
}

# Aggregate over `hr_bin` factor.
data <- data %>%
    select(-hr) %>%
    group_by(dteday, hr_bin) %>%
    summarize(
        holiday = mode(holiday),
        temp = mean(temp),
        atemp = mean(atemp),
        windspeed = mean(atemp),
        cnt = sum(cnt)
    )

head(data, 10)
```

### Adding dummy variables
We factorise the categorical data present within the data set. A method to do this for regression analysis is by
creating dummy variables (a.k.a. dummy coding). By removing the first dummy variable - leaving us with n-1 dummy
variables - we avoid multi-collinearity issues in  (https://cran.r-project.org/web/packages/fastDummies/fastDummies.pdf).\
More info on categorical data in regression (https://stats.oarc.ucla.edu/spss/faq/coding-systems-for-categorical-variables-in-regression-analysis-2/#:~:text=Categorical%20variables%20require%20special%20attention,entered%20into%20the%20regression%20model.) \ 

(Further research: generalised linear model, using poisson distribution.)
```{r dummy}
data <- dummy_cols(data, select_columns = c("hr_bin"), remove_first_dummy = FALSE)
head(data, 10)
```