---
title: "Bike Sharing"
author: "Aga, Karla, Nisse, Ole"
date: "`r Sys.Date()`"
output: html_document
---

## Load data

Data set is acquired from [machine-learning repo](https://archive.ics.uci.edu/dataset/275/bike+sharing+dataset).

Import the required dependencies.

```{r import, message = FALSE}
library(tidyverse)
library(fastDummies)
library(kableExtra)
library(gridExtra, exclude="combine")
library(lubridate)
```

Read the data.

```{r read bike}
source_data <- read.csv("../data/hour.csv", header = TRUE, sep = ",") %>%
  as_tibble()

data <- source_data
```

Expand date data with `# day` and date numeric conversion.
```{r expand date}
# Converts datetime string to numeric with origin offset.
date_to_num <- function(date, offset) {
   as.integer(as.Date(date, "%Y-%m-%d")) - offset
}

# Fetch origin.
dt_offset <- source_data$dteday[1] %>%
  date_to_num(0) - 1


data <- data %>%
  mutate(
    day = day(dteday),
    dt_num = date_to_num(dteday, dt_offset)
  )

# Global variable for amount of unique days.
n_days <- n_distinct(data$dteday)
```

Sub-select variables.

```{r select}
data <- select(data, 
               dt_num,
               yr,
               day,
               mnth,
               hr, 
               weathersit, 
               temp, 
               atemp, 
               hum, 
               windspeed, 
               cnt)
```

Summary of our data.

```{r summary}
summary(data)
head(data)
```
Data contains ``r n_days`` unique days, meaning that the entirety of 2011 (365 days) and 2012 (366 days) was recorded.

## Preprocessing & EDA

### Missing data

Check standard NA values. There are none.

```{r missing standard}
anyNA(data)
```

### Distributions

Distributions of variables.

```{r distributions, message = FALSE, echo = FALSE}
grid.arrange(
    ggplot(data, aes(hr)) + geom_histogram(binwidth = 1),
    ggplot(data, aes(weathersit)) + geom_histogram(binwidth = 1),
    ggplot(data, aes(temp)) + geom_histogram(),
    ggplot(data, aes(atemp)) + geom_histogram(),
    ggplot(data, aes(hum)) + geom_histogram(),
    ggplot(data, aes(windspeed)) + geom_histogram(),
    ggplot(data, aes(cnt)) + geom_histogram()
)
```

#### Missing records for hour

Some `hr` values seem to have a lower amount of occurrences than others. 

```{r hr missing}
hr_df <- data %>%
  group_by(hr) %>%
  summarise(
    n = n(),
    n_missing_days = n_days - n)

hr_df
```

Determine where missing records are located.

```{r missing location}
missing_df <- data %>%
  group_by(dt_num) %>%
  summarise(
    n_records = n(),
    n_missing_records = 24 - n_records
  ) %>%
  filter(n_missing_records > 0) +
  arrange(desc(n_missing_records))

missing_df
```
